---
title: "EDA"
author: "Taylor Bonar & Satvik Ajmera"
date: "2022-07-24"
output: powerpoint_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r}
library(tswge)
library(arrow)

btc = read_parquet("data/btc_3month_2021.parquet")
head(btc)
```

```{r}
summary(btc)
```
```{r}
plot(btc$timestamp, btc$Target, type="l",
     main="Bitcoin Residualized Returns in March-May 2021",
     xlab="Date",
     ylab="15 Minute Residualized Returns")
```

```{r}
plotts.sample.wge(btc$Target)

end.idx = length(btc$Target)
mid.idx = end.idx/2
midplusone.idx = mid.idx + 1

acf(btc$Target[1:mid.idx],plot=T, ylim=c(-1,1))
acf(btc$Target[midplusone.idx:end.idx],plot=T, ylim=c(-1,1))
```
***Stationarity Check:**
  * Mean - appears dependent on time
  * Variance - appears to not be constant
  * Correlations - follow similar patterns, however anomalies can be observed.
  
  
We have 7 peaks of frequency, with the highest near 0 (about 0.0165). This in turns means we have a frequency of 0.0165 and cycle of around 60.

Let's try filtering out some of the dataset to focus on particular behaviors of interest
```{r}
ma = filter(btc$Target, rep(1,5)/5)
plot(ma, type="l")
```

```{r}
ma = filter(btc$Target, rep(1,60)/60)
plot(ma, type="l")
```

```{r}
ma = filter(btc$Target, rep(1,1440)/1440)
plot(ma, type="l")
```

```{r}
pw = parzen.wge(btc$Target)

peak.pzgram.index = which.max(pw$pzgram)

pw$freq[1]

```
# Multiple Regression w/ Correlated Errors
```{r}
library(GGally)
ggpairs(btc[3:10])
```

Analyzing Close Prices on BTC
```{r}
plot(btc$timestamp, btc$Close, type = "l",
     main="BTC's Close Price over Period",
     xlab="Date",
     ylab="Close Price")
```

# Hypothetical BTC Target Data
```{r}
plot(btc$timestamp, btc$Count, type="l",
     main="BTC Count over Mar-May 2021",
     xlab="Date",
     ylab="Count")
```

```{r}
plot(btc$timestamp, btc$Open, type="l",
     main="BTC Open Price over Mar-May 2021",
     xlab="Date",
     ylab="Opening Price")
```

```{r}
plot(btc$timestamp, btc$High, type="l",
     main="BTC High over Mar-May 2021",
     xlab="Date",
     ylab="Highest Price")
```

```{r}
plot(btc$timestamp, btc$Low, type="l",
     main="BTC Low over Mar-May 2021",
     xlab="Date",
     ylab="Lowest Price")
```

```{r}
plot(btc$timestamp, btc$Close, type="l",
     main="BTC Close over Mar-May 2021",
     xlab="Date",
     ylab="Close Price")
```

```{r}
plot(btc$timestamp, btc$Volume, type="l",
     main="BTC Volume over Mar-May 2021",
     xlab="Date",
     ylab="Number of CryptoAsset Units Traded")
```

```{r}
plot(btc$timestamp, btc$VWAP, type="l",
     main="BTC VWAP over Mar-May 2021",
     xlab="Date",
     ylab="Volume WEighted Average Price")
```
# Univariate Analysis

We'll take a look at two univariate analysis, but as we saw earlier, 5 out of 6 of our features are serially correlated. We'll use one from that grouping of 5 and do a second analysis on Volume.

## Open Univariate Analysis

```{r}
plotts.sample.wge(btc$Open)
```
```{r}
end.idx = length(btc$Open)
mid.idx = end.idx/2
midplusone.idx = mid.idx + 1

acf(btc$Open[1:mid.idx],plot=T, ylim=c(-1,1))
acf(btc$Open[midplusone.idx:end.idx],plot=T, ylim=c(-1,1))
```
Transform to remove non-stationarity
Find the cycle period:
```{r}
pw_open = parzen.wge(btc$Open)

peak.pzgram.index = which.max(pw_open$pzgram)

pw_open$freq[1]

```
Spectral density shows a peaked frequency of 0.00000758 giving a cycle period of 131,926 minutes.

**Hourly Transformation**
```{r}
btc_trans_open_h = artrans.wge(btc$Open, c(rep(0,59),1))
```
```{r}
aic5.wge(btc_trans_open_h)
```

```{r}
ljung.wge(btc_trans_open_h, p=2, q=2)
ljung.wge(btc_trans_open_h, p=2, q=2, K=48)
```

**Daily Transformation**
```{r}
btc_trans_open_d = artrans.wge(btc$Open, c(rep(0,1439),1))
```
```{r}
aic5.wge(btc_trans_open_d)
```

```{r}
ljung.wge(btc_trans_open_d, p=1, q=1)
ljung.wge(btc_trans_open_d, p=1, q=1, K=48)
```

**Weekly Transformation**
```{r}
btc_trans_open_w = artrans.wge(btc$Open, c(rep(1,10079),1))
```

```{r}
aic5.wge(btc_trans_open_w)
```

```{r}
ljung.wge(btc_trans_open_w, p=3, q=0)
ljung.wge(btc_trans_open_w, p=3, q=0, K=48)
```
We confidently reject that any of the Open feature is not white noise. We will need to use methods that are non-stationary to properly model the realization's behavior.

## Predictions
**Hourly**
```{r}
predsPart_open_h = fore.aruma.wge(btc$Open, s = 60, n.ahead = 1440, lastn=TRUE)
```
```{r}
plot(predsPart_open_h$f, type = "l")
plot(seq(1,132000,1),
     btc$Open,
     type = "l",
     ylab = "BTC Open Prices",
     main = "24 Hours BTC Open Forecast")
lines(seq(130560,132000,1), predsPart_open_h$f, type = "l", col = "red")
```

## Volume Univariate Analysis
```{r}
plotts.sample.wge(btc$Volume)

end.idx = length(btc$Volume)
mid.idx = end.idx/2
midplusone.idx = mid.idx + 1

acf(btc$Volume[1:mid.idx],plot=T, ylim=c(-1,1))
acf(btc$Volume[midplusone.idx:end.idx],plot=T, ylim=c(-1,1))
```

Transform to remove non-stationarity
```{r}
btc_trans_volume = artrans.wge(btc$Volume, c(rep(1,20000),1))
```

```{r}
aic5.wge(btc$Volume)
```

# Multivariate Analysis

**Step 1:** Perform a regression analysis and model the residuals.
```{r}
ksfit = lm(Target~Count+Open+High+Low+Close+Volume+VWAP, data=btc)

phi = aic.wge(ksfit$residuals, p=0:20, q=0)
```

**Step 2:** Use ARIMA to perform the MLE analysis which estimates the coefficients in the multiple regression while simultaneously modeling Z_t as an AR(phi$p)
```{r}
fit = arima(btc$Target, order=c(phi$p, 0,0), xreg=btc[,3:9])
fit
AIC(fit)
```
With a model fitted to not account for lags or trends, we can see the following:
* Our model's coefficients are very small (under 10) with some exceptions:
  * AR(1) w/ a t-stat of 806.75
  
* Our psuedo-t-statistic (coefficient/s.e.)


We fitted a model with no lags or trends, we'll take the fit$resid and check the residual plots and/or a Ljung-Box Test to confirm white noise.
```{r}
acf(fit$residuals)
ltest = ljung.wge(fit$resid)
ltest$pval
```

Visually we can see that our Residuals for ACF are not consistent with white noise. Additionally our ljung test shows a p-value of 0.

Let's add a time trend!
```{r}
ksfit2 = lm(Target~timestamp+Count+Open+High+Low+Close+Volume+VWAP, data=btc)
phi2 = aic.wge(ksfit2$residuals, p=0:8, q=0)
```

```{r}
fit2 = arima(btc$Target, order=c(phi2$p, 0,0), xreg=cbind(btc$unix_time, btc[,3:9]))
fit2
```






