---
title: "EDA"
author: "Taylor Bonar & Satvik Ajmera"
date: "2022-07-24"
output: powerpoint_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r}
library(tswge)
library(arrow)

btc = read_parquet("data/btc_3month_2021.parquet")
head(btc)
pw = parzen.wge(btc$Target)

peak.pzgram.index = which.max(pw$pzgram)

pw$freq[1]
plotts.wge(btc$Target)

```
```{r}
pw$freq[1]
```


```{r}
pw.VWAP = parzen.wge(btc$VWAP)

peak.pzgram.index = which.max(pw.VWAP$pzgram)

pw.VWAP$freq[1]
```

```{r}
pw.VWAP = parzen.wge(btc$VWAP)

peak.pzgram.index = which.max(pw.VWAP$pzgram)

pw.VWAP$freq[1]
```

```{r}
parzen.wge(btc$Count)
```

```{r}
parzen.wge(btc$Open)
```


```{r}
parzen.wge(btc$Volume)

```

```{r}
acf(btc$Volume,lag.max = 10000)
```

```{r}
acf(btc$Target,lag.max = 50)
```

```{r}
acf(btc$Open,lag.max = 100000)
22000

```



```{r}
acf(btc$Volume,lag.max = 100000)
```


```{r}
parzen.wge(btc$Volume)
```

```{r}
#forecast Particles
plotts.sample.wge(CM$part) #freq near .0192 (annual)
btc_20000 = artrans.wge(btc$Volume, c(rep(0,19999),1))

plotts.sample.wge(btc_20000) #looks like some low freq?
aic5.wge(CM_52) #picks ARMA(2,1) assume stationary
aic5.wge(CM_52,type = "bic") #picks ARMA(0,0) 
ljung.wge(CM_52)$pval #FTR Ho
ljung.wge(CM_52, K = 48)$pval #FTR Ho
#Going with white noise despite peak at 0 in Spec D. 
#est = est.arma.wge(CM_52, p = 3, q = 2)
#CM_52_AR2_MA1 = artrans.wge(CM_52,est$phi)
predsPart = fore.aruma.wge(CM$part,s = 52, n.ahead = 20)
plot(predsPart$f, type = "l")
plot(seq(1,508,1), CM$part, type = "l",xlim = c(0,528), ylab = "Temperature", main = "20 Week Particulate Forecast")
lines(seq(509,528,1), predsPart$f, type = "l", col = "red")

```


```{r}

```


```{r}
#Hourly Transformation
btc_trans_open_h = artrans.wge(btc$Volume, c(rep(0,59),1))
```
Daily Transformation
```{r}
btc_trans_open_d = artrans.wge(btc$Volume, c(rep(0,1439),1))
```
Weekly Transformation
```{r}
btc_trans_open_w = artrans.wge(btc$Volume, c(rep(0,10079),1))
```

```{r}
aic5.wge(btc$Volume)
```

```{r}
aic5.wge(btc_trans_open_h)
```

```{r}
aic5.wge(btc_trans_open_d)
```

```{r}
aic5.wge(btc_trans_open_w)
```


```{r}
predsPart_open_h = fore.aruma.wge(btc_trans_open_h, s = 60, n.ahead = 1440, lastn=TRUE)


```


```{r}
plot(predsPart_open_h$f, type = "l")
plot(seq(1,120480,1), btc$Open, type = "l",xlim = c(0,121920), ylab = "BTC Open Prices", main = "24 Hours BTC Open Forecast")
lines(seq(120481,121920,1), predsPart$f, type = "l", col = "red")
```

```{r}
plot(predsPart$f, type = "l")
plot(seq(1,508,1), CM$part, type = "l",xlim = c(0,121920), ylab = "Temperature", main = "20 Week Particulate Forecast")
lines(seq(509,121920,1), predsPart$f, type = "l", col = "red")
```


```{r}
summary(btc)
```
```{r}
plot(btc$timestamp, btc$Target, type="l",
     main="Bitcoin Residualized Returns in January 2018",
     xlab="Date",
     ylab="15 Minute Residualized Returns")
```
***Stationarity Check:**
  * Mean - appears dependent on time
  * Variance - appears to not be constant
  * Correlations - follow similar patterns, however anomalies can be observed.

```{r}
plotts.sample.wge(btc$Target)

end.idx = length(btc$Target)
mid.idx = end.idx/2
midplusone.idx = mid.idx + 1

acf(btc$Target[1:mid.idx],plot=T, ylim=c(-1,1))
acf(btc$Target[midplusone.idx:end.idx],plot=T, ylim=c(-1,1))
```

We have 7 peaks of frequency, with the highest near 0. Let's try filtering out some of the dataset to focus on particular behaviors of interest
```{r}
ma = filter(btc$Target, rep(1,5)/5)
plot(ma, type="l")
```

```{r}
ma = filter(btc$Target, rep(1,60)/60)
plot(ma, type="l")
```


```{r}
CMsmall = CM[1:478,]
```


Forecast 1440 minutes ahead
```{r}
btcSMALL.short = btc[1:120480,]

ksfit = lm(Target~Open+Volume, data = btcSMALL.short)
ksfit
phi = aic.wge(ksfit$residuals)
phi

fit = arima((btcSMALL.short,order = c(phi$p,phi$q,0), period = 1440), xreg = cbind=(btcSMALL.short$Open,btcSMALL.short$Volume))
```


Forecast 10080
```{r}

```




#Find ASE  Need to forecast last 30 of known series.  
CMsmall = CM[1:478,]
ksfit = lm(cmort~temp+part+Week+FWeek, data = CMsmall)
phi = aic.wge(ksfit$residuals)
fit = arima(CMsmall$cmort,order = c(phi$p,0,0), seasonal = list(order = c(1,0,0), period = 52), xreg = cbind(CMsmall$temp, CMsmall$part, CMsmall$Week, CMsmall$FWeek))
AIC(fit) #AIC = 2991.85

last30 = data.frame(temp = CM$temp[479:508], part = CM$part[479:508], Week = seq(479,508,1), FWeek = as.factor(seq(479,508,1)%%52))
#get predictions
predsCMort = predict(fit,newxreg = last30)

#predict residuals manually
plotts.sample.wge(ksfit$residuals)
phi = aic.wge(ksfit$residuals)
resids = fore.arma.wge(ksfit$residuals,phi = phi$phi,n.ahead = 30)
#predict trend manually
preds = predict(ksfit, newdata = last30)

predsFinal = preds + resids$f


plot(seq(1,508,1), CM$cmort, type = "l",xlim = c(0,528), ylab = "Cardiac Mortality", main = "20 Week Cardiac Mortality Forecast")
lines(seq(479,508,1), predsFinal, type = "l", col = "red")


ASE = mean((CM$cmort[479:508] - predsFinal)^2,na.rm = TRUE)
ASE #32.2167



