---
title: "univar_Taylor"
author: "Taylor Bonar"
date: "2022-07-27"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(lubridate)  # Unix to Date Timestamps/POSIXct
library(arrow)
library(tswge)
library(tidyverse)

train = read_parquet("data/btc-train.parquet")
test.short = read_parquet("data/btc-test-short.parquet")
test.long = read_parquet("data/btc-test-long.parquet")
```
# Data Cleanup
```{r}
summary(train)
```

```{r}
summary(test.short)
```

```{r}
summary(test.long)
```

Converting UNIX timestamps to POSIXct so they are human readable
```{r}
train <- train %>% mutate(timestamp = as_datetime(timestamp))
test.short <- test.short %>% mutate(timestamp = as_datetime(timestamp))
test.long <- test.long %>% mutate(timestamp = as_datetime(timestamp))
```

# EDA Exploration
```{r}
library(GGally)

ggpairs(train[3:9])
```
We'll use Open as our response, and try building a univariate model or two with Volume and Count

Let's first look at our time series plots to get an idea of how Open operates
```{r}
library(tswge)

plotts.wge(train$Open)
plotts.sample.wge(train$Open)
```
We can see from the plotts.wge that we have a wandering realization of the Open variable for BTC. We most likely have a positive phi to estimate.

# Univariate Models
 
```{r}
aic5.wge(train$Open, p=0:10, q=0:10) # p=9, q=9
aic5.wge(train$Open, p=0:10, q=0:10, type="bic") # p=10, q=0
```
## Model 1 -- AIC
### ARMA(9,9)
```{r}
aic.est.arma = est.arma.wge(train$Open, p=9, q=9)
aic.est.arma$phi
aic.est.arma$theta
aic.est.arma$avar
```
#### Short-Term Forecast
```{r}
forecast.period = 1440
aic.fore.short.arma = fore.arma.wge(train$Open, phi=aic.est.arma$phi, theta = aic.est.arma$theta, n.ahead=forecast.period, lastn=TRUE, limits=TRUE)
```
```{r}
end.idx = length(train$Open)
start.idx = end.idx - forecast.period+1

aic.short.ASE = mean((train$Open[start.idx:end.idx] - aic.fore.short.arma$f)^2)
aic.short.ASE
# ASE = 66,8584.3
sqrt(aic.long.ASE)
# Root ASE = 817.67

roll.win.rmse.wge(train$Open, horizon = forecast.period, s = 0, d = 0, phi=aic.est.arma$phi, theta = aic.est.arma$theta)c
# Rolling Window RMSE = 1333.199
```
```{r}
forecast.period = 1440
aic.fore.short.arma = fore.arma.wge(train$Open, phi=aic.est.arma$phi, theta = aic.est.arma$theta, n.ahead=forecast.period, lastn=FALSE, limits=TRUE)
```

#### Long-Term Forecast

```{r}
forecast.period = 10080
aic.fore.long.arma = fore.arma.wge(train$Open, phi=aic.est.arma$phi, theta = aic.est.arma$theta, n.ahead=forecast.period, lastn=TRUE, limits=TRUE)
```

```{r}
end.idx = length(train$Open)
start.idx = end.idx - forecast.period+1

aic.long.ASE = mean((train$Open[start.idx:end.idx] - aic.fore.short.arma$f)^2)
aic.long.ASE
# ASE = 4073064
sqrt(aic.long.ASE)
# Root ASE = 2018.183


roll.win.rmse.wge(train$Open, horizon = forecast.period, s = 0, d = 0, phi=aic.est.arma$phi, theta = aic.est.arma$theta)

# Rolling Window RMSE = 3585.325
```
```{r}
forecast.period = 10080
aic.fore.long.arma = fore.arma.wge(train$Open, phi=aic.est.arma$phi, theta = aic.est.arma$theta, n.ahead=forecast.period, lastn=FALSE, limits=TRUE)
```


## Model 2 -- BIC
### ARMA(10,0)
```{r}
bic.est.arma = est.ar.wge(train$Open, p=10)
bic.est.arma$phi
bic.est.arma$theta
bic.est.arma$avar
```
```{r}
forecast.period = 1440
bic.fore.short.ar = fore.arma.wge(train$Open, phi=bic.est.arma$phi, theta=0, n.ahead=forecast.period, lastn=TRUE, limits=TRUE)
```

```{r}
end.idx = length(train$Open)
start.idx = end.idx - forecast.period+1

bic.short.ASE = mean((train$Open[start.idx:end.idx] - bic.fore.short.ar$f)^2)
bic.short.ASE
# ASE 703111.0144

roll.win.rmse.wge(train$Open, horizon = 1440, s = 0, d = 0, phi=bic.est.arma$phi, theta = 0)

# Rolling Window RMSE is:  1332.593
```
```{r}
forecast.period = 1440
bic.fore.short.ar = fore.arma.wge(train$Open, phi=bic.est.arma$phi, theta=0, n.ahead=forecast.period, lastn=FALSE, limits=TRUE)
```
#### Long-term Forecast

```{r}
forecast.period = 10080
bic.fore.long.ar = fore.arma.wge(train$Open, phi=bic.est.arma$phi, theta=0, n.ahead=forecast.period, lastn=TRUE, limits=TRUE)
```

```{r}
end.idx = length(train$Open)
start.idx = end.idx - forecast.period+1

bic.long.ASE = mean((train$Open[start.idx:end.idx] - bic.fore.long.ar$f)^2)
bic.long.ASE
# ASE = 17879034
sqrt(bic.long.ASE)
# Root ASE = 4228.361
```

```{r}
roll.win.rmse.wge(train$Open, horizon = forecast.period, s = 0, d = 0, phi=bic.est.arma$phi, theta = 0)

# Rolling Window RMSE = 3580.7
```

```{r}
forecast.period = 10080
bic.fore.long.ar = fore.arma.wge(train$Open, phi=bic.est.arma$phi, theta=0, n.ahead=forecast.period, lastn=FALSE, limits=TRUE)
```



